---
title: "Cormack-Jolly-Seber Model with Dipper Data"
output: html_document
---

---

\   

### Dipper Cormack-Jolly-Seber (CJS) Model

The dipper model is a very common example CJS model.  As used, for example, in <a href="https://link.springer.com/article/10.1007/s10651-016-0353-z" target="_blank" style="color: blue">Turek et al. (2016)</a>, *Environmental and Ecological Statistics*.

The observational data consists binary 1's and 0's, representing observed (1), or not-observed (0).  The data is for $N$ = 209 birds, over a period of $T$ = 7 years.

Of the $N$ = 209 birds, 101 are females, and 108 are males.  Initially we'll ignore gender, for simplicity, but we'll use the gender information later.

The parameters of interest will be survival $\phi$, and probability of detection $p$.  Both of these are assumed to be constant in our first model.

\   

### Dipper Data

You can <a href="https://danielturek.github.io/nimble-cefe-2019/data/dipper_data.Rdata" style="color: blue">download the dipper data here</a>.

This file contains the variables:

- `sightings`: An $N$ = 209 by $T$ = 7 array, containing 1's, 0's and `NA`'s.  Each row represents the sighting history of a unique bird.  Each column is a different year.  Leading `NA` values are observation periods before an individual was first observed.
- `first`: A vector of length $N$ = 209, where each entry gives the time period for which an individual was first seen.  For example, individal #1 was first seen on time period 1, therefore `first[1] = 1`, whereas individual #13 was first seen on time period 2, therefore `first[13] = 2`.
- `gender`: A vector of length $N$ = 209, where each entry specifies the gender of a individual.  1's correspond to females, and 2's correspond to males.

\  

### Fitting the dipper model

Let's construct model code for the dipper CJS model.

Then, we'll fit the model using NIMBLE's MCMC.

In our model, we'll use the following nodes:

- `phi`: Probability of survival
- `p`: Probability detection (capture) probability
- `x[i, t]`: Latent states representing true alive (1) or dead (0) status, for individual `i` at time `t`
- `y[i, t]` = Observed data for capture (1) or not-captured (0) for individual `i` at time `t`
- `first[i]` = Year of first capture for individual `i`

We'll also have to take care to provide appropriate initial values for the `x` matrix.


\   

### Results

```{r, echo = FALSE, message = FALSE}
library(nimble, quiet = TRUE)
load('~/github/nimble/nimble-cefe-2019/docs/data/dipper_samples.Rdata')
```

Here are the results I got, for our first dipper CJS model:

```{r }
round(samplesSummary(samples), 3)
```







\  

\  

\  

\  

\  

\  

\  

\  

\  

\  

\  

\  

\  

\  





