---
title: "Comparing MCMC Performance: `compareMCMCs`"
output: html_document
---

---

\   


### Package `compareMCMCs`

NIMBLE provides a nice tool for comparing MCMC performance, called `compareMCMCs`.

This actually used to be part of the core NIMBLE package.  Now, it's on the way to its own package, which will be called `compareMCMCs`.

Right now, you can install the (future) `compareMCMCs` package from GitHub, using this code:

\   

```{r eval = FALSE}
library(devtools)

devtools::install_github("nimble-dev/compareMCMCs", subdir = "compareMCMCs")

library(compareMCMCs)
```

\   


The general framework for using `compareMCMCs` is:

```{r eval = FALSE}
comp <- compareMCMCs(modelInfo,
                     MCMCs,
                     nimbleMCMCdefs,
                     monitors,
                     MCMCcontrol)

make_MCMC_comparison_pages(comp, dir, modelName)
```

\   

Let's try using `compareMCMCs`, to compare these different versions of the litters model.



\  

\  

\  

\  

\  

\  

\  

\  

\  

\  

\  

\  

\  

\  


