---
title: "One-Line MCMC: `nimbleMCMC`"
output: html_document
---

---

\   

By popular request, NIMBLE also provides a "one-liner" for MCMC.

The entire process of

- Creating model
- Creating MCMC
- Compiling
- Running the MCMC

can be done using the `nimbleMCMC` function.

We'll use our old friend the dipper model, as an example:

```{r eval = FALSE}
library(nimble)

load('~/Downloads/dipper_data.Rdata')

dipperCode <- nimbleCode({
    phi ~ dunif(0, 1)
    p ~ dunif(0, 1)
    for(i in 1:N) {
        x[i, first[i]] <- 1
        y[i, first[i]] <- 1
        for(t in (first[i]+1):T) {
            x[i, t] ~ dbern(phi * x[i, t-1])
            y[i, t] ~ dbern(p * x[i, t])
        }
    }
})

dipperConsts <- list(N = dim(sightings)[1], T = dim(sightings)[2], first = first)

dipperData <- list(y = sightings)

xInit <- ifelse(!is.na(y), 1, 0)

dipperInits <- list(phi = 0.5, p = 0.5, x = xInit)
```

\   

### One-line invocation of MCMC: `nimbleMCMC`

Now, we'll use `nimbleMCMC` to fit the dipper model:

```{r eval = FALSE}
samples <- nimbleMCMC(dipperCode, dipperConsts, dipperData, dipperInits)
```

\   

`nimbleMCMC` provides a variety of arguments:

```{r eval = FALSE}
out <- nimbleMCMC(dipperCode, dipperConsts, dipperData, dipperInits,
                  nchains = 3,
                  samples = FALSE,
                  summary = TRUE,
                  WAIC = TRUE)
```

You can find more information using `help(nimbleMCMC)`.



\  

\  

\  

\  

\  

\  

\  

\  

\  

\  

\  

\  

\  

\  



