---
title: "Multi-State Capture-Recapture Model: Comparisons"
output: html_document
---

---

\   

### Comparing the orchid model results

The `compareMCMCs` package that we saw before, also lets you compare the results from **different** models, if they have top-level parameters in common.

We'll use this, to compare the results from the two different versions of the orchids model.


\   

```{r echo = FALSE}
load('~/github/nimble/nimble-cefe-2019/docs/data/orchid_samples.Rdata')
load('~/github/nimble/nimble-cefe-2019/docs/data/orchidHMM_samples.Rdata')
```



### Rename the MCMCs

Both versions of the MCMC were called `"nimble"`.  So we'll rename them to better represent the results: `latent`, and `HMM`.

The function `renameMCMC` takes the output from `compareMCMCs`, and the **new** name for an MCMC, then the **old** name.

```{r eval = FALSE}
library(compareMCMCs)

out <- renameMCMC(out, 'latent', 'nimble')

outHMM <- renameMCMC(outHMM, 'HMM', 'nimble')
```

\  

### Combine the MCMC results together

This is done using simple list concatenation, using R's `c()` function.

```{r eval = FALSE}
outCombined <- c(out, outHMM)
```

\   

### Make the comparison results

```{r eval = FALSE}
make_MCMC_comparison_pages(outCombined, dir = '~/Desktop/orchid', modelName = 'orchid')
```

\  

The results <a href="https://danielturek.github.io/nimble-cefe-2019/fig/orchid/orchid" target="_blank" style="color: blue">are
available here</a>.











\  

\  

\  

\  

\  

\  

\  

\  

\  

\  

\  

\  

\  

\  




